Dismal Code: Studying the Evolution of Security Bugs

### 2013-04-08 docs/evol.tex
reverted to hypothesis

A security bug is a programming error that introduces a potentially
exploitable weakness into a computer system~\cite{SSL12}\cite{TJBD11}\cite{M06}. This weakness could lead to a
security breach with unfortunate consequences in different layers, like databases,
native code, applications, libraries and others. Despite the significant
effort to detect and eliminate such bugs~\cite{SZ12}, little attention has been paid to
study them in relation to software evolution~\cite{L96}\cite{LRWPT97}\cite{IB06}\cite{RGMA06}.
In this paper we present how we used a large software ecosystem to analyse
how different types of security vulnerabilities are related to the evolving software packages.

One of the most common approaches to identify security bugs is
{\it static analysis}~\cite{CW07}. This kind of analysis involves the
inspection of the program's source or object code without executing
it. For our research we used {\it FindBugs},\footnote{\url{http://findbugs.sourceforge.net/}}
a static analysis tool that examines bytecode to detect software bugs and has already been used in
research~\cite{AP10}\cite{HP07}\cite{HP04}\cite{HW08}\cite{SHP06}.
Specifically, we ran FindBugs on all the project
versions of all the projects that exist in the
{\it Maven Central Repository}\footnote{\url{http://search.maven.org/}}
(approximately 265GB of data---see Section~\ref{sec:data}).
Then we observed the changes that involved the security bugs and their characteristics.
This research builds upon our earlier work on the topic~\cite{MGS12}.

We chose to focus our study on the security bugs rather than other
software bugs. This is because compared to other bug categories,
the category of security bugs has two distinct features: it is critical~\cite{SZ12}
and it costs~\cite{BCL08}\c~\cite{TH04}ite{R06}. Specifically, a software bug can
lead to a malfunction of a software artifact that runs under specific
requirements but a security bug can allow a malicious user to alter the execution
of the entire application for his or her own gain. In this case, such bugs could span a wide
range of security and privacy issues, like viewing sensitive information, the destruction or
modification of sensitive data, denial of service and others.
Moreover, security bug disclosures lead to a negative and significant change
in market value for a software vendor~\cite{TW07}.
Hence, one of the basic pursuits in every new software release should
be the mitigation of such bugs.

The motivation behind our work was to validate if programmers tend to care for
such high-risk bugs when they release a new version of their software. In
addition, we wanted to investigate other critical features associated with such
vulnerabilities like the persistence of a bug; in essence, to check if critical bugs stay
unresolved for a long time. Also, we wanted to elaborate more on the relation of security
bugs not only with performance bugs (as Zaman et al.~\cite{ZAH11} have already
done) but also with other bug categories. In the same manner, we tried to check
the relation between version size and security bugs, knowing the
that research has been contradictory on the the
issue~\cite{BP84}\cite{SYTP85}\cite{NBZ06}\cite{GKMS00}.
Finally, we examined the Maven ecosystem as a whole from a security
perspective. Its structure gave us the opportunity to see if a project version that is a dependency to
a large number of others contains a low rate of security bugs~\cite{MW10}.
%In this case we can strengthen the relationship between Linus' Law and
%software security as Meneely et al. have already indicated~\cite{MW10}.

The main contributions of this research are:
\begin{itemize}
	\item The analysis of how security bugs found through static analysis,
evolve over time. To achieve this, we inspect every project per version.
Our hypothesis is that security bugs should decrease as a project evolves since
they are critical and developers should eliminate them.
	\item Security bug persistence between versions. We expect that
security bugs should be eliminated earlier than other bugs.
	\item The relation between security bugs and a project
version's size. Our hypothesis is that security bugs are proportional to a project
version's size.  Here, size is defined in terms of bytecode.
	\item The correlation of security bugs with other bug categories. Our
hypothesis is that security bugs flock together with bugs that have to do with
performance, coding practices and others.
%	\item The relation of software dependencies and security bugs. Our
%hypothesis is that an artifact that is a dependency of many other artifacts
%should have a low rate of security bugs. 
\end{itemize}

@@

Our primary hypothesis was that security bugs and especially severe
ones, should be corrected as projects evolve. However, our findigs indicate that this is
not true. Contrary to our second research hypothesis, we showed that security
bugs are not eliminated in a way that is particularly different from the other bugs.
Also, having an average of two to three versions persistence in a sample where 60\%
of the projects have three versions, is not a positive result especially in the
case of the {\it Security High} bugs. Concerning the relation between severe security bugs and a project's size
we showed evidence that they are not proportionally related. Furthermore, the
pairwise correlations between all categories indicated that even though all the other
categories are related, severe bugs do not flock together with the other bugs.
Also, it is interesting to see that security bugs were one of the top two
bug categories existing in a large ecosystem. Finally, we highlighted the
domino effect, and showed evidence that indicates that the Linus' Law does not
apply in the case of the security bugs.

>>>>

A security bug is a programming error that introduces a potentially exploitable
weakness into a computer system~\cite{SSL12}\cite{TJBD11}\cite{M06}. This
weakness could lead to a security breach with unfortunate consequences in
different layers, like databases, native code, applications, libraries and
others. Despite the significant effort to detect and eliminate such
bugs~\cite{SZ12}, little attention has been paid to study them in relation to
software evolution~\cite{L96}\cite{LRWPT97}\cite{IB06}\cite{RGMA06}. 
One of the most common approaches to identify security bugs is
{\it static analysis}~\cite{CW07}. This kind of analysis involves the
inspection of the program's source or object code without executing
it.

In this
paper we present how we used a large software ecosystem to analyse how different
types of security vulnerabilities are related to the evolving software packages.
For our research we used {\it FindBugs},\footnote{\url{http://findbugs.sourceforge.net/}}
a static analysis tool that examines bytecode to detect software bugs and has already been used in
research~\cite{AP10}\cite{HP07}\cite{HP04}\cite{HW08}\cite{SHP06}.
Specifically, we ran FindBugs on all the project
versions of all the projects that exist in the
{\it Maven Central Repository}\footnote{\url{http://search.maven.org/}}
(approximately 265GB of data---see Section~\ref{sec:data}).
Then we observed the changes that involved the security bugs and their characteristics.
This research builds upon our earlier work on the topic~\cite{MGS12}.

We chose to focus our study on security bugs rather than other types of
software defects. This is because compared to other bug categories,
the category of security bugs has two distinct features: it is critical~\cite{SZ12}
for an organization's infrastructure
and moreover it can cause significant financial damage to 
an organization~\cite{TH04}\cite{BCL08}\cite{R06}. 
Specifically, a software bug can
lead to a malfunction of a software artifact that runs under specific
requirements but a security bug can allow a malicious user to alter the execution
of the entire application for his or her own gain. In this case, such bugs could span a wide
range of security and privacy issues, like viewing sensitive information, the destruction or
modification of sensitive data, denial of service and others.
Moreover, security bug disclosures lead to a negative and significant change
in market value for a software vendor~\cite{TW07}.
Hence, one of the basic pursuits in every new software release should
be to mitigate such bugs.

% @dimitro: say why we did this work, not how we did it
The motivation behind our work was to validate whether programmers care for
such high-risk bugs when they release a new version of their software. In
addition, we wanted to investigate other critical features associated with such
vulnerabilities like the persistence of a bug; in essence, to check if critical bugs stay
unresolved for a long time. Also, we wanted to elaborate more on the relation of security
bugs not only with performance bugs (as Zaman et al.~\cite{ZAH11} have already
done) but also with other bug categories. In the same manner, we tried to check
the relation between version size and security bugs, knowing the
that research has been contradictory on the the
issue~\cite{BP84}\cite{SYTP85}\cite{NBZ06}\cite{GKMS00}.
Finally, we examined the Maven ecosystem as a whole from a security
perspective. Its structure gave us the opportunity to see if a project version that is a dependency to
a large number of others contains a low rate of security bugs~\cite{MW10}.
%In this case we can strengthen the relationship between Linus' Law and
%software security as Meneely et al. have already indicated~\cite{MW10}.

In this work we:
\begin{itemize}
	\item Analyze how security bugs found through static analysis,
evolve over time. To achieve this, we inspect every project per version.
Our hypothesis is that security bugs should decrease as a project evolves since
they are critical and developers should eliminate them.
%We present evidence that defect counts remain relatively stable
%across project revisions.
	\item Examine security bug persistence between versions. 
We expect that security bugs should be eliminated earlier than other bugs.
%We find that a specific defect remains unfixed on average 
%for 3 revisions.
	\item Study the relation between security bugs and a project
version's size. Our hypothesis is that security bugs are proportional to a project
version's size. Here, size is defined in terms of bytecode.
%We show that many defect types are correlated with the project's
%size (with the notable exception of high-risk security defects).
	\item Examine the correlation of security bugs with other bug categories.
Our hypothesis is that security bugs flock together with bugs that
have to do with performance, coding practices and others.
%We show that a code base can have defect hotspots.
%	\item The relation of software dependencies and security bugs. Our
%hypothesis is that an artifact that is a dependency of many other artifacts
%should have a low rate of security bugs. 
  %\item Make available a carefully constructed dataset and a statistical toolkit for
   % performing analysis of the Java Maven ecosystem.
\end{itemize}

@@

Our primary hypothesis was that security bugs and especially severe
ones, should be corrected as projects evolve. However, our findigs indicate that this is
not true. Contrary to our second research hypothesis, we showed that security
bugs are not eliminated in a way that is particularly different from the other bugs.
Also, having an average of two to three versions persistence in a sample where 60\%
of the projects have three versions, is not a positive result especially in the
case of the {\it Security High} bugs. Concerning the relation between severe security bugs and a project's size
we showed evidence that they are not proportionally related. Furthermore, the
pairwise correlations between all categories indicated that even though all the other
categories are related, severe bugs do not flock together with the other bugs.
Also, it is interesting to see that security bugs were one of the top two
bug Make available a carefully constructed dataset and a statistical toolkit for
    performing analysis of the Java Maven ecosystem.
categories existing in a large ecosystem. Finally, we highlighted the
domino effect, and showed evidence that indicates that the Linus' Law does not
apply in the case of the security bugs.
